---
title: "Untitled"
author: "Peiran Zhang"
date: "2022/7/6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG="ENG.UK")
rm(list = ls())
gc()
```

Loading the required libraries.
```{r}
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(dplyr)
library(bnlearn)
library(Rgraphviz)
library(MASS)
```

## Section 3 (Exploratory Data Analysis)
```{r, fig.width=5}
## load the whole easySHARE dataset and extract those in wave 5
load("easySHARE.rda")
wave5 = easySHARE_rel8_0_0[easySHARE_rel8_0_0$wave==5,]

## we only want to study the following variables
vbs = c("female","age", "isced1997_r", "hhsize", "partnerinhh", "mother_alive", "father_alive", "ch001_", "ch021_mod", "sp002_mod", "chronic_mod", "eurod", "br010_mod", "adla", "iadlza","mobilityind", "bmi2", "br015_", "smoking", "recall_1", "recall_2", "numeracy_1", "numeracy_2")
wave5 = wave5[vbs]

## we focus on those older than or at 60 years old
wave5 = wave5[wave5$age >= 60,]

## print the summary of current data
summary(wave5)

## deal with unrealistic number of grandchildren (Section 3.1)
wave5$ch021_mod = ifelse((wave5$ch021_mod/wave5$ch001_)>17,-99,wave5$ch021_mod)

################### Section 3.2 ###################
## deal with missing values 
# loop to replace the -9 with 0, and other negative value with NA
for (i in 1:length(vbs)) {
  v.name = vbs[i]
  mid.var = wave5[v.name]
  mid.var[mid.var == -9] = 0
  mid.var[mid.var < 0] = NA
  wave5[v.name] = mid.var
}

################### Section 3.3 ###################
## some binary variables have strange codes which need to be converted
wave5$partnerinhh[wave5$partnerinhh==3] = 0
wave5$mother_alive[wave5$mother_alive==5] = 0
wave5$father_alive[wave5$fathe_alive==5] = 0
wave5$sp002_mod[wave5$sp002_mod==5] = 0
wave5$smoking[wave5$smoking==5] = 0

################### Section 3.4 ###################
## create a new variable, soc_sup, to represent the social support situation
wave5$soc_con = wave5$hhsize + wave5$partnerinhh +
  wave5$mother_alive + wave5$father_alive + wave5$ch001_ + wave5$ch021_mod + wave5$sp002_mod


## drop other social support variables
wave5 = subset(wave5, select = -c(hhsize, partnerinhh, 
                  mother_alive, father_alive,
                   ch001_, ch021_mod, sp002_mod))

## create a new variable, numeracy, combining two numeracy scores
numeracy = rep(NA,dim(wave5)[1])
# if numeracy_1 and numeracy_2 are not missing, get their average values
numeracy[!is.na(wave5$numeracy_1) & !is.na(wave5$numeracy_2)] = (wave5$numeracy_1[!is.na(wave5$numeracy_1) & !is.na(wave5$numeracy_2)] + wave5$numeracy_2[!is.na(wave5$numeracy_1) & !is.na(wave5$numeracy_2)] )/2
# if one of the variables are missing, use the left ones
numeracy[!is.na(wave5$numeracy_1) & is.na(wave5$numeracy_2)] = wave5$numeracy_1[!is.na(wave5$numeracy_1) & is.na(wave5$numeracy_2)]
numeracy[is.na(wave5$numeracy_1) & !is.na(wave5$numeracy_2)] = wave5$numeracy_2[is.na(wave5$numeracy_1) & !is.na(wave5$numeracy_2)]

wave5$numeracy = numeracy

wave5 = subset(wave5, select = -c(numeracy_1, numeracy_2))

## create a new variable, cogscore, to represent the cognitive function status
cogscore = rep(NA,dim(wave5)[1])
cogscore = wave5$recall_1 + wave5$recall_2 + wave5$numeracy
wave5$cogscore = cogscore

## create a new variable, funscore, to represent the ability of performing daily physical tasks
funscore = rep(NA,dim(wave5)[1])
funscore = wave5$adla + wave5$iadlza + wave5$mobilityind
wave5$funscore = funscore

## rename some variables
wave5 <- wave5 %>% rename(education = isced1997_r)
wave5 <- wave5 %>% rename(drinking = br010_mod)
wave5 <- wave5 %>% rename(do_act = br015_)

## turn some variables into factors directly
wave5$female = factor(wave5$female)
wave5$education = factor(wave5$education)
wave5$drinking = factor(wave5$drinking)
wave5$bmi2 = factor(wave5$bmi2, ordered = T)
wave5$do_act = factor(wave5$do_act)
wave5$smoking = factor(wave5$smoking, ordered = T)

################### Section 3.4 ###################
## some bar plots for categorical variables (Figure 1)
pg1 = ggplot() + 
  geom_bar(aes(x = wave5$female),color="darkblue", fill="lightblue")+
  labs(x ="Gender")
pe1 = ggplot() + 
  geom_bar(aes(x = wave5$education),color="darkblue", fill="lightblue")+
  labs(x ="Education")
pdb1 = ggplot() + 
  geom_bar(aes(x = wave5$drinking),color="darkblue", fill="lightblue")+
  labs(x ="Drinking behavior")
pb1 = ggplot() + 
  geom_bar(aes(x = wave5$bmi2),color="darkblue", fill="lightblue")+
  labs(x ="BMI type")
pda1 = ggplot() + 
  geom_bar(aes(x = wave5$do_act),color="darkblue", fill="lightblue")+
  labs(x ="Doing activities")
ps1 = ggplot() + 
  geom_bar(aes(x = wave5$smoking),color="darkblue", fill="lightblue")+
  labs(x ="Smoking")
grid.arrange(pg1, pe1, pdb1, pb1, pda1, ps1, nrow = 2)


## convert drinking into an ordered factor and combine some levels
wave5$drinking[wave5$drinking ==2 | wave5$drinking ==3] =2
wave5$drinking[wave5$drinking ==4 | wave5$drinking ==5] =3
wave5$drinking[wave5$drinking ==6 | wave5$drinking ==7] =4
wave5$drinking = factor(wave5$drinking,order = T)

## convert education into an ordered factor and combine some levels, combine 1 and 2 into basic education, 3 and 4 into secondary education, 5 and 6 into tertiary education, others into NA
wave5$education[wave5$education ==1 | wave5$education ==2] =1
wave5$education[wave5$education ==3 | wave5$education ==4] =2
wave5$education[wave5$education ==5 | wave5$education ==6] =3
wave5$education[wave5$education ==97 | wave5$education ==95] =NA
wave5$education = factor(wave5$education,order = T)

## due to the imbalanced distribution of do_act, we combine level 2 and 3 as doing sports seldom 
wave5$do_act <- ifelse(wave5$do_act == 3, 2, wave5$do_act)
wave5$do_act <- ifelse(wave5$do_act == 4, 3, wave5$do_act)
wave5$do_act <- factor(wave5$do_act, ordered = T)

## some bar plots for discrete variables (Figure 2)
pg2 = ggplot() + 
  geom_bar(aes(x = wave5$female),color="darkblue", fill="lightblue")+
  labs(x ="Gender")
pe2 = ggplot() + 
  geom_bar(aes(x = wave5$education),color="darkblue", fill="lightblue")+
  labs(x ="Education")
pdb2 = ggplot() + 
  geom_bar(aes(x = wave5$drinking),color="darkblue", fill="lightblue")+
  labs(x ="Drinking behavior")
pb2 = ggplot() + 
  geom_bar(aes(x = wave5$bmi2),color="darkblue", fill="lightblue")+
  labs(x ="BMI type")
pda2 = ggplot() + 
  geom_bar(aes(x = wave5$do_act),color="darkblue", fill="lightblue")+
  labs(x ="Doing activities")
ps2 = ggplot() + 
  geom_bar(aes(x = wave5$smoking),color="darkblue", fill="lightblue")+
  labs(x ="Smoking")
grid.arrange(pg2, pe2, pdb2, pb2, pda2, ps2, nrow = 2)

################### Section 3.6 ###################
## histograms for continuous variables (Figure 3)
pa1 = ggplot() +
  geom_histogram(aes(x=wave5$age), color="darkblue", fill="lightblue") +
  labs(x ="Age")

pcs1 = ggplot() +
  geom_histogram(aes(x=wave5$cogscore), color="darkblue", fill="lightblue") +
  labs(x ="Cognitive score")

## some bar plots for discrete variables (Figure 3)
pcd1 = ggplot() +
  geom_bar(aes(x=wave5$chronic_mod), color="darkblue", fill="lightblue") + 
labs(x ="Chronic disease")
pds1 = ggplot() +
  geom_bar(aes(x=wave5$eurod), color="darkblue", fill="lightblue") +
  labs(x ="Depression scale")
psc1 = ggplot() +
  geom_bar(aes(x=wave5$soc_con), color="darkblue", fill="lightblue") +
  labs(x ="Social contact")
pfs1 = ggplot() +
  geom_bar(aes(x=wave5$funscore), color="darkblue", fill="lightblue") +
  labs(x ="Function score")

grid.arrange(pa1, pcd1, pds1, psc1, pcs1, pfs1, nrow = 2)

## due to the imbalanced distribution of eurod, we combine all observations greater than 4 as level 4, and make the variable an ordered factor
wave5$eurod <- ifelse(wave5$eurod > 4, 4, wave5$eurod)
wave5$eurod <- factor(wave5$eurod, order = T)
# rename
wave5 <- wave5 %>% rename(depression = eurod)

## due to the imbalanced distribution of chronic_mod, We set three levels for the number of chronic diseases: 0 (healthy), 1 (one disease) and 2 (more than one diseases)
wave5$chronic_mod <- ifelse(wave5$chronic_mod > 2, 2, wave5$chronic_mod)
wave5$chronic_mod <- factor(wave5$chronic_mod, order = T)

## print the summary of soc_con
summary(wave5$soc_con)
# due to the imbalanced distribution of soc_con, we set 2 levels for it: 0 (little social contact) and 1 (enough social contact)
# the threshold is the median, 12
wave5$soc_con <- ifelse(wave5$soc_con > 12, 1, 0)
wave5$soc_con <- factor(wave5$soc_con, ordered = T)

## due to the imbalanced distribution of funscore, we set 2 levels: 0 (normal) and 1 (difficult)
wave5$funscore <- ifelse(wave5$funscore == 0, 0, 1)
wave5$funscore <- factor(wave5$funscore, ordered = T)

################### Section 3.7 ###################

## violin plots for categorical variables and cognitive score (Figure 4)

# index of observations that are not missing in the corresponding variables
inde <- !is.na(wave5$education)
inddb <- !is.na(wave5$drinking)
indb <- !is.na(wave5$bmi2)
indda <- !is.na(wave5$do_act)
inds <- !is.na(wave5$smoking)

pg3 <- ggplot(wave5, aes(x=female, y=cogscore, fill=female)) + 
  geom_violin(trim = FALSE) + labs(x ="Gender")
pe3 <- ggplot(wave5[inde,], aes(x=education, y=cogscore, fill=education)) +
  geom_violin(trim = FALSE) + labs(x ="Education")
pdb3 <- ggplot(wave5[inddb,], aes(x=drinking, y=cogscore, fill=drinking)) + 
  geom_violin(trim = FALSE) + labs(x ="Drinking behavior")
pb3 <- ggplot(wave5[indb,], aes(x=bmi2, y=cogscore, fill=bmi2)) + 
  geom_violin(trim = FALSE) + labs(x ="BMI type")
pda3 <- ggplot(wave5[indda,], aes(x=do_act, y=cogscore, fill=do_act)) + 
  geom_violin(trim = FALSE) + labs(x ="Doing activities")
ps3 <- ggplot(wave5[inds,], aes(x=smoking, y=cogscore, fill=smoking)) + 
  geom_violin(trim = FALSE) + labs(x ="Smoking")
grid.arrange(pg3, pe3, pdb3, pb3, pda3, ps3, nrow = 2)

## next, we convert continuous variables into factors
wave5$age <- cut(wave5$age, breaks = c(seq(60,85,5),105), ordered_result = TRUE, right = FALSE)
wave5$cogscore <- cut(wave5$cogscore, breaks = c(0,quantile(wave5$cogscore, probs = c(.1,.4),na.rm=TRUE),max(wave5$cogscore, na.rm = TRUE)+1), 
                         ordered_result = TRUE, right = FALSE)

## plot some stacked bar plots for new factor variables
# the index of observations that are not missing for the corresponding variables
indcs <- !is.na(wave5$cogscore)
indcc <- !is.na(wave5$chronic_mod) & !is.na(wave5$cogscore)
indds <- !is.na(wave5$depression) & !is.na(wave5$cogscore)
indsc <- !is.na(wave5$soc_con) & !is.na(wave5$cogscore)
indfs <- !is.na(wave5$funscore) & !is.na(wave5$cogscore)
# stacked bar plots
pa2.1 = ggplot(wave5[indcs,], aes(x = age, fill = cogscore)) + 
  geom_bar() + labs(x ="Age")
pa2.2 = ggplot(wave5[indcs,], aes(x = age, fill = cogscore)) + 
  geom_bar(position = "fill") + labs(x ="Age", y = "proportion")
pcd2.1 = ggplot(wave5[indcc,], aes(x = chronic_mod, fill = cogscore)) + 
  geom_bar() + labs(x ="Chronic disease")
pcd2.2 = ggplot(wave5[indcc,], aes(x = chronic_mod, fill = cogscore)) + 
  geom_bar(position = "fill") + labs(x ="Chronic disease",  y = "proportion")
pds2.1 = ggplot(wave5[indds,], aes(x = depression, fill = cogscore)) + 
  geom_bar() + labs(x ="Depression")
pds2.2 = ggplot(wave5[indds,], aes(x = depression, fill = cogscore)) + 
  geom_bar(position = "fill") + labs(x ="Depression",  y = "proportion")
psc2.1 = ggplot(wave5[indsc,], aes(x = soc_con, fill = cogscore)) + 
  geom_bar() + labs(x ="Social contact")
psc2.2 = ggplot(wave5[indsc,], aes(x = soc_con, fill = cogscore)) + 
  geom_bar(position = "fill") + labs(x ="Social contact",  y = "proportion")
pfs2.1 = ggplot(wave5[indfs,], aes(x = funscore, fill = cogscore)) + 
  geom_bar() + labs(x ="Funtion score")
pfs2.2 = ggplot(wave5[indfs,], aes(x = funscore, fill = cogscore)) + 
  geom_bar(position = "fill") + labs(x ="Funtion score",  y = "proportion")
grid.arrange(pa2.1, pa2.2, pcd2.1, pcd2.2, pds2.1, pds2.2, psc2.1, psc2.2, pfs2.1, pfs2.2, nrow = 5)

## drop some redundant variables
wave5.fin = subset(wave5, select = -c(adla, iadlza, 
                  mobilityind, recall_1,
                   recall_2, numeracy))

```





## Bayesian network code (Section 5.1)
```{r, fig.width=5}
## black list
myblacklist = matrix(c(
                   "female","age",
                   "education","age",
                   "smoking","age",
                   "chronic_mod","age",
                   "drinking","age",
                   "bmi2","age",
                   "do_act","age",
                   "soc_con","age",
                   "funscore","age",
                   "cogscore","age",
                   "age","female",
                   "education","female",
                   "smoking","female",
                   "chronic_mod","female",
                   "drinking","female",
                   "bmi2","female",
                   "do_act","female",
                   "soc_con","female",
                   "funscore","female",                   
                   "cogscore","female",
                   "smoking","education",
                   "chronic_mod","education",
                   "drinking","education",
                   "bmi2","education",
                   "do_act","education",
                   "soc_con","education",
                   "funscore","education",                         
                   "cogscore","education",
                   "bmi2", "smoking",
                   "funscore", "chronic_mod"), 
                 byrow = TRUE, ncol=2, dimnames =list(NULL,c("from","to")))
## ci tests
ci.test("female", "education", test = "mi", data = wave5.fin) #p-value < 2.2e-16
ci.test("soc_con", "depression", test = "jt", data = wave5.fin) #p-value = 0.9729
ci.test("smoking", "depression", test = "jt", data = wave5.fin) #p-value = 0.01956
ci.test("smoking", "cogscore", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("do_act", "bmi2", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("drinking", "bmi2", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("soc_con", "female", test = "mi", data = wave5.fin) #p-value = 2.253e-05
ci.test("depression", "cogscore", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("smoking", "chronic_mod", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("do_act", "chronic_mod", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("bmi2", "cogscore", test = "jt", data = wave5.fin) #p-value = 0.1071
ci.test("age", "funscore", test = "jt", data = wave5.fin) #p-value < 2.2e-16
ci.test("female", "do_act", test = "mi", data = wave5.fin) #p-value < 2.2e-16
ci.test("age", "do_act", test = "jt", data = wave5.fin) #p-value < 2.2e-16
## white list
mywhitelist = matrix(c(
                   "female", "education",
                   "smoking", "depression",
                   "do_act", "bmi2",
                   "drinking", "bmi2",
                   "female", "soc_con",
                   "depression", "cogscore",
                   "smoking", "chronic_mod",
                   "do_act", "chronic_mod",
                   "age", "funscore",
                   "female", "do_act",
                   "age", "education",
                   "age", "do_act",
                   "female", "cogscore",
                   "do_act", "funscore",
                   "age", "cogscore",
                   "age","soc_con",
                   "female", "smoking"
               ), 
                 byrow = TRUE, ncol=2, dimnames =list(NULL,c("from","to")))
## learn structure of DAG
wave5.iamb =iamb(wave5.fin, whitelist = mywhitelist,blacklist = myblacklist)
graphviz.plot(wave5.iamb)
bn.bayes = bn.fit(wave5.iamb, data=wave5.fin, method = "bayes", iss = 10)
```

## Local distributions (Section 5.2)

First, we need to use custom.fit to learn the local distributions via ordered logistic regression. 
```{r}
## check the dimensions of the probability matrix and corresponding variables
dim(bn.bayes$cogscore$prob)
# 3:cogscore 2:female 6:age 4: education 5:depression 4:drinking
```
Fit the model of cognitive score using ordered logistic regression.
```{r}
# fir for female
olr.fitf =polr(cogscore ~ age + education + depression + drinking, data = wave5.fin[wave5.fin$female==1,])
# fit for male
olr.fitm =polr(cogscore ~ age + education + depression + drinking, data = wave5.fin[wave5.fin$female==0,])
# extract levels of each variable
edu.lv = levels(wave5.fin$education)
age.lv = levels(wave5.fin$age)
f.lv = levels(wave5.fin$female)
d.lv = levels(wave5.fin$depression)
dr.lv = levels(wave5.fin$drinking)
cs.lv = levels(wave5.fin$cogscore)
# create a data frame having all combinations of all independent variables
combos = data.frame(
age=rep(age.lv, length(edu.lv)),
education=rep(edu.lv,each = length(age.lv)),
depression=rep(d.lv, each =
length(age.lv)*length(edu.lv)),
drinking=rep(dr.lv, each =
length(age.lv)*length(edu.lv)*length(d.lv))
)
# predict the probabilities in every situation
distcsf = predict(olr.fitf, newdata= combos ,type="p")
distcsm = predict(olr.fitm, newdata= combos ,type="p")
# convert the results above into the format of bn.fit
distcs = array(dim = c(length(cs.lv), length(f.lv), length(age.lv), length(edu.lv), length(d.lv), length(dr.lv)), dimnames = list(cogscore = cs.lv, female = f.lv,  age = age.lv, education = edu.lv, depression = d.lv, drinking = dr.lv))
distcs[,1,,,,] = array(t(distcsm), dim=c(length(cs.lv),1, length(age.lv),  length(edu.lv), length(d.lv), length(dr.lv)))
distcs[,2,,,,] = array(t(distcsf), dim=c(length(cs.lv),1, length(age.lv),  length(edu.lv), length(d.lv), length(dr.lv)))
# custom fit finished
bn.bayes2 = bn.bayes
bn.bayes2$cogscore = distcs
```


## Plot of cognitive score and education (Section 5.2.1)
```{r,fig.width=5}
p1 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,1,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, No education, No depression, No drinking")


p2 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,1,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, No education, No depression, No drinking")

p3 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, No depression, No drinking")

p4 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education, No depression, No drinking")
p5 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,3,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Secondary education, No depression, No drinking")

p6 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,3,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Secondary education, No depression, No drinking")
p7 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,4,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Tertiary education, No depression, No drinking")

p8 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,4,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Tertiary education, No depression, No drinking")

grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8,  nrow = 4)
```
Below is the calculations of the improvement of the probabilities of having high cognitive score for males.

The improvement for females, from no education to basic education, in all age groups
```{r}
(bn.bayes2$cogscore$prob[,2,,2,1,1][3,] - bn.bayes2$cogscore$prob[,2,,1,1,1][3,])/bn.bayes2$cogscore$prob[,2,,1,1,1][3,]
```

The improvement for males, from no education to basic education, in all age groups
```{r}
(bn.bayes2$cogscore$prob[,1,,2,1,1][3,] - bn.bayes2$cogscore$prob[,1,,1,1,1][3,])/bn.bayes2$cogscore$prob[,1,,1,1,1][3,]
```

The improvement for females, from basic education to secondary education, in all age groups
```{r}
(bn.bayes2$cogscore$prob[,2,,3,1,1][3,] - bn.bayes2$cogscore$prob[,2,,2,1,1][3,])/bn.bayes2$cogscore$prob[,2,,2,1,1][3,]
```

The improvement for males, from basic education to secondary education, in all age groups
```{r}
(bn.bayes2$cogscore$prob[,1,,3,1,1][3,] - bn.bayes2$cogscore$prob[,1,,2,1,1][3,])/bn.bayes2$cogscore$prob[,1,,2,1,1][3,]
```

The improvement for females, from secondary education to tertiary education, in all age groups
```{r}
(bn.bayes2$cogscore$prob[,2,,4,1,1][3,] - bn.bayes2$cogscore$prob[,2,,3,1,1][3,])/bn.bayes2$cogscore$prob[,2,,3,1,1][3,]
```

The improvement for males, from basic education to secondary education, in all age groups
```{r}
(bn.bayes2$cogscore$prob[,1,,4,1,1][3,] - bn.bayes2$cogscore$prob[,1,,3,1,1][3,])/bn.bayes2$cogscore$prob[,1,,3,1,1][3,]
```
 
The probabilities of getting low cognitive score for men aged 85 years and older in the selected depression level and drinking behavior groups 
```{r}
bn.bayes2$cogscore$prob[,1,,1,1,1][1,6] #0.47011
bn.bayes2$cogscore$prob[,1,,2,1,1][1,6] #0.2269042
bn.bayes2$cogscore$prob[,1,,3,1,1][1,6] #0.1148473
bn.bayes2$cogscore$prob[,1,,4,1,1][1,6] #0.05910975
```


## Plot of cognitive score and depression (Section 5.2.2)
```{r,fig.width=7}
p9 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, No Depression, No drinking")


p10 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education, No Depression, No drinking")

p11 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,2,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, Slight Depression, No drinking")

p12 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,2,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,Slight Depression, No drinking")
p13 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,3,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, Mild Depression, No drinking")

p14 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,3,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,Mild Depression, No drinking")
p15 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,4,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, Moderate Depression, No drinking")

p16 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,4,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,Moderate Depression, No drinking")
p17 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,5,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, Severe Depression, No drinking")

p18 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,5,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,Severe Depression, No drinking")
grid.arrange(p9, p10, p11, p12, p13, p14, p15, p16, p17, p18, nrow = 5)
```

In all five depression levels, for males aged 85 years and older,
the probabilities of getting low cognitive score are 18.60%/51.86%/115.45%/167.09%/364.91% higher than those of having high cognitive score.
```{r}
(bn.bayes2$cogscore$prob[,1,,2,1,1][1,6]-bn.bayes2$cogscore$prob[,1,,2,1,1][3,6])/bn.bayes2$cogscore$prob[,1,,2,1,1][3,6] #0.1860001
(bn.bayes2$cogscore$prob[,1,,2,2,1][1,6]-bn.bayes2$cogscore$prob[,1,,2,2,1][3,6])/bn.bayes2$cogscore$prob[,1,,2,2,1][3,6] #0.5186284
(bn.bayes2$cogscore$prob[,1,,2,3,1][1,6]-bn.bayes2$cogscore$prob[,1,,2,3,1][3,6])/bn.bayes2$cogscore$prob[,1,,2,3,1][3,6] #1.154539
(bn.bayes2$cogscore$prob[,1,,2,4,1][1,6]-bn.bayes2$cogscore$prob[,1,,2,4,1][3,6])/bn.bayes2$cogscore$prob[,1,,2,4,1][3,6] #1.670899
(bn.bayes2$cogscore$prob[,1,,2,5,1][1,6]-bn.bayes2$cogscore$prob[,1,,2,5,1][3,6])/bn.bayes2$cogscore$prob[,1,,2,5,1][3,6] #3.649124
```

The same situation appears among female respondents, but the numbers are slightly smaller: 11.25%/37.32%/51.12%/92.79%/289.00%.

```{r}
(bn.bayes2$cogscore$prob[,2,,2,1,1][1,6]-bn.bayes2$cogscore$prob[,2,,2,1,1][3,6])/bn.bayes2$cogscore$prob[,2,,2,1,1][3,6] #0.1124881
(bn.bayes2$cogscore$prob[,2,,2,2,1][1,6]-bn.bayes2$cogscore$prob[,2,,2,2,1][3,6])/bn.bayes2$cogscore$prob[,2,,2,2,1][3,6] #0.3731571
(bn.bayes2$cogscore$prob[,2,,2,3,1][1,6]-bn.bayes2$cogscore$prob[,2,,2,3,1][3,6])/bn.bayes2$cogscore$prob[,2,,2,3,1][3,6] #0.5111639
(bn.bayes2$cogscore$prob[,2,,2,4,1][1,6]-bn.bayes2$cogscore$prob[,2,,2,4,1][3,6])/bn.bayes2$cogscore$prob[,2,,2,4,1][3,6] #0.9279015
(bn.bayes2$cogscore$prob[,2,,2,5,1][1,6]-bn.bayes2$cogscore$prob[,2,,2,5,1][3,6])/bn.bayes2$cogscore$prob[,2,,2,5,1][3,6] #2.890041
```


The probabilities of getting low cognitive score are 94.10%/92.73%/90.11%/87.98%/79.57% lower than those of having high cognitive score among men aged between 60 and 65years in five depression levels.
```{r}
(bn.bayes2$cogscore$prob[,1,,2,1,1][1,1]-bn.bayes2$cogscore$prob[,1,,2,1,1][3,1])/bn.bayes2$cogscore$prob[,1,,2,1,1][3,1] #-0.9410056
(bn.bayes2$cogscore$prob[,1,,2,2,1][1,1]-bn.bayes2$cogscore$prob[,1,,2,2,1][3,1])/bn.bayes2$cogscore$prob[,1,,2,2,1][3,1] #-0.9272565
(bn.bayes2$cogscore$prob[,1,,2,3,1][1,1]-bn.bayes2$cogscore$prob[,1,,2,3,1][3,1])/bn.bayes2$cogscore$prob[,1,,2,3,1][3,1] #-0.9011477
(bn.bayes2$cogscore$prob[,1,,2,4,1][1,1]-bn.bayes2$cogscore$prob[,1,,2,4,1][3,1])/bn.bayes2$cogscore$prob[,1,,2,4,1][3,1] #-0.8798459
(bn.bayes2$cogscore$prob[,1,,2,5,1][1,1]-bn.bayes2$cogscore$prob[,1,,2,5,1][3,1])/bn.bayes2$cogscore$prob[,1,,2,5,1][3,1] #-0.7956939
```


## Plot of cognitive score and drinking behavior (Section 5.2.3)
```{r,fig.width=5}
p17 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, No Depression, No Drinking")


p18 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,1,1]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education, No Depression, No Drinking")

p19 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,1,2]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, No Depression, Rarely")

p20 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,1,2]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,No Depression, Rarely")
p21 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,1,3]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, No Depression, Sometimes")

p22 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,1,3]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,No Depression, Sometimes")
p23 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,1,,2,1,4]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Male, Basic education, No Depression, Often")

p24 = ggplot(mapping = aes(x = rep(seq(1:length(levels(wave5.fin$age))),3),
                     y=matrix(t(bn.bayes2$cogscore$prob[,2,,2,1,4]), ncol =1), color = rep(levels(wave5.fin$cogscore), each = length(levels(wave5.fin$age))))) + 
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks=seq(1:length(levels(wave5.fin$age))),
        labels=levels(wave5.fin$age)) +
  labs(x = "Age", y= "Conditional probability", color = "Cognitive score", title ="Female, Basic education,No Depression, Often")
grid.arrange(p17, p18, p19, p20, p21, p22, p23, p24, nrow = 4)
```
Taking women aged between 60 and 65 years in the selected groups as example, the probabilities of getting high cognitive score are 0.68/0.79/0.80/0.77 for drinking behavior 1/2/3/4 respectively.
```{r}
bn.bayes2$cogscore$prob[,2,,2,1,1][3,1] #0.6825861
bn.bayes2$cogscore$prob[,2,,2,1,2][3,1] #0.7946212
bn.bayes2$cogscore$prob[,2,,2,1,3][3,1] #0.8043686
bn.bayes2$cogscore$prob[,2,,2,1,4][3,1] #0.771251
```

Likewise, these numbers are 0.64/0.72/0.75/0.72 for men with the same characteristics.
```{r}
bn.bayes2$cogscore$prob[,1,,2,1,1][3,1] #0.6392302
bn.bayes2$cogscore$prob[,1,,2,1,2][3,1] #0.7187197
bn.bayes2$cogscore$prob[,1,,2,1,3][3,1] #0.7512143
bn.bayes2$cogscore$prob[,1,,2,1,4][3,1] #0.7158799
```

In our selected groups, women aged between 60 and 65 years in all four drinking behavior levels are 1490.18%/3167.10%/3409.50%/2673.23% more likely to have high cognitive score than having low cognitive score.
```{r}
(bn.bayes2$cogscore$prob[,2,,2,1,1][3,1]-bn.bayes2$cogscore$prob[,2,,2,1,1][1,1])/bn.bayes2$cogscore$prob[,2,,2,1,1][1,1] #14.90182
(bn.bayes2$cogscore$prob[,2,,2,1,2][3,1]-bn.bayes2$cogscore$prob[,2,,2,1,2][1,1])/bn.bayes2$cogscore$prob[,2,,2,1,2][1,1] #31.67096
(bn.bayes2$cogscore$prob[,2,,2,1,3][3,1]-bn.bayes2$cogscore$prob[,2,,2,1,3][1,1])/bn.bayes2$cogscore$prob[,2,,2,1,3][1,1] #34.09499
(bn.bayes2$cogscore$prob[,2,,2,1,4][3,1]-bn.bayes2$cogscore$prob[,2,,2,1,4][1,1])/bn.bayes2$cogscore$prob[,2,,2,1,4][1,1] #26.73226

```

In contrast, the numbers for women aged between 80 and 85 years are 109.89%/397.52%/442.48%/308.16%.
```{r}
(bn.bayes2$cogscore$prob[,2,,2,1,1][3,5]-bn.bayes2$cogscore$prob[,2,,2,1,1][1,5])/bn.bayes2$cogscore$prob[,2,,2,1,1][1,5] #1.098911
(bn.bayes2$cogscore$prob[,2,,2,1,2][3,5]-bn.bayes2$cogscore$prob[,2,,2,1,2][1,5])/bn.bayes2$cogscore$prob[,2,,2,1,2][1,5] #3.975147
(bn.bayes2$cogscore$prob[,2,,2,1,3][3,5]-bn.bayes2$cogscore$prob[,2,,2,1,3][1,5])/bn.bayes2$cogscore$prob[,2,,2,1,3][1,5] #4.424765
(bn.bayes2$cogscore$prob[,2,,2,1,4][3,5]-bn.bayes2$cogscore$prob[,2,,2,1,4][1,5])/bn.bayes2$cogscore$prob[,2,,2,1,4][1,5] #3.081549
```